---
title: "Quarto Reference"
subtitle: "Syntax cheatsheet for HTML documents"
author: "Emile Latour"
date: "today"
author-title: "By"
published-title: "Date"
toc: true
toc_float:
  collapsed: false
format:
  html:
    link-external-newwindow: true
    toc-location: left
    number-sections: true
    number-depth: 4
    code-tools: true
    code-fold: true
    self-contained: true
    embed-resources: true
    linkcolor: blue
execute:
  echo: true
  eval: false
editor_options:
  chunk_output_type: console
---

> **How to use this doc:** This is a living reference, not a template to fill in.
> Copy the snippets you need, paste them into your analysis, and delete the rest.
> Render this doc to see each element in action.

```{r}
#| label: setup
#| include: false
#| eval: true

library(tidyverse)
library(palmerpenguins)
```


# Callout Boxes

Five types are available: `note`, `tip`, `warning`, `caution`, and `important`.

## Default appearance

::: {.callout-note}
This is a **note** callout. Good for supplementary information.
:::

::: {.callout-tip}
This is a **tip** callout. Good for helpful suggestions.
:::

::: {.callout-warning}
This is a **warning** callout. Good for things the reader should watch out for.
:::

::: {.callout-caution}
This is a **caution** callout. Similar to warning but slightly softer.
:::

::: {.callout-important}
This is an **important** callout. Use sparingly.
:::

## With a custom title

::: {.callout-note}
## A custom title goes here

Body text goes here. The `##` heading inside the div sets the title.
:::

## Collapsible (collapsed by default)

::: {.callout-caution collapse="true"}
## Click to expand

This content is hidden until the reader expands it. Useful for lengthy
asides, proofs, or supplementary detail you don't want to dominate the page.
:::

## Collapsible (expanded by default)

::: {.callout-tip collapse="false"}
## Expanded by default, but collapsible

The reader can collapse this if they want.
:::

## Appearance variants

```
appearance = "default"   # icon + coloured left border + shaded background
appearance = "simple"    # icon + coloured left border, no shaded background  
appearance = "minimal"   # coloured left border only, no icon
```

::: {.callout-note appearance="simple"}
## Simple appearance

No shaded background.
:::

::: {.callout-note appearance="minimal"}
## Minimal appearance

Just the left border — very low visual weight.
:::

## Without icon

::: {.callout-note icon=false}
## No icon

The icon is suppressed with `icon=false`.
:::


# Tabsets

Tabsets let you show alternative content (e.g. plot vs table, or Model A vs B)
without cluttering the page.

## Basic tabset

::: {.panel-tabset}

### Tab one

Content for the first tab. Can be prose, code output, a plot — anything.

### Tab two

Content for the second tab.

### Tab three

Content for the third tab.

:::

## Tabset with code and plot

::: {.panel-tabset}

### Plot

```{r}
#| eval: true
#| echo: false

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, colour = species)) +
  geom_point(alpha = 0.7) +
  labs(title = "Penguin size", x = "Flipper length (mm)", y = "Body mass (g)") +
  theme_minimal()
```

### Data

```{r}
#| eval: true
#| echo: false

penguins |>
  head(10) |>
  knitr::kable()
```

### Code

```{r}
#| eval: false
#| echo: true

ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, colour = species)) +
  geom_point(alpha = 0.7) +
  labs(title = "Penguin size", x = "Flipper length (mm)", y = "Body mass (g)") +
  theme_minimal()
```

:::


# Columns

## Two equal columns

::: {.columns}

::: {.column width="50%"}
**Left column**

Any content here — prose, a list, a plot, a table.
:::

::: {.column width="50%"}
**Right column**

Content here.
:::

:::

## Unequal columns

::: {.columns}

::: {.column width="30%"}
**Narrow sidebar**

Good for a legend, a note, or a key takeaway.
:::

::: {.column width="5%"}
<!-- spacer -->
:::

::: {.column width="65%"}
**Wider main content**

The 5% spacer column adds breathing room between the two.
:::

:::

## Three columns

::: {.columns}

::: {.column width="33%"}
Column 1
:::

::: {.column width="33%"}
Column 2
:::

::: {.column width="33%"}
Column 3
:::

:::


# Font Size

Scale text up or down inline using a `div` with a `style` attribute.

::: {style="font-size: 150%;"}
This text is 150% of normal size.
:::

::: {style="font-size: 75%;"}
This text is 75% of normal size. Good for fine print or captions under custom figures.
:::

Common values: `50%`, `75%`, `90%`, `110%`, `125%`, `150%`, `200%`.


# Code Chunk Annotation

Add numbered callouts to lines of code with `#<N>` comments.
The annotations appear below the chunk.

```{r}
#| eval: true

set.seed(20240101)        #<1>
x <- rnorm(100, mean = 0) #<2>
mean(x)                   #<3>
```

1. Set the seed for reproducibility.
2. Draw 100 samples from a standard normal distribution.
3. Compute and print the sample mean.


# Figure Layout

## Side-by-side figures

```{r}
#| eval: true
#| echo: false
#| layout-ncol: 2
#| fig-cap:
#|   - "Flipper length vs body mass"
#|   - "Flipper length vs bill length"

ggplot(penguins, aes(flipper_length_mm, body_mass_g, colour = species)) +
  geom_point(alpha = 0.7) + theme_minimal() + theme(legend.position = "bottom")

ggplot(penguins, aes(flipper_length_mm, bill_length_mm, colour = species)) +
  geom_point(alpha = 0.7) + theme_minimal() + theme(legend.position = "bottom")
```

## Figure with caption and cross-reference

```{r}
#| label: fig-penguins
#| fig-cap: "Body mass versus flipper length for three penguin species."
#| eval: true
#| echo: false

ggplot(penguins, aes(flipper_length_mm, body_mass_g, colour = species)) +
  geom_point(alpha = 0.7) +
  theme_minimal()
```

Then refer to it in prose as `@fig-penguins` → @fig-penguins.


# Inline Code Values

Pull values from R directly into prose so numbers never get out of sync.

```{r}
#| eval: true
#| include: false

n_penguins <- nrow(penguins)
mean_mass  <- round(mean(penguins$body_mass_g, na.rm = TRUE))
```

> The dataset contains `r n_penguins` penguins with a mean body mass of
> `r scales::comma(mean_mass)` g.


# Useful YAML Options (quick reference)

```yaml
format:
  html:
    toc: true                    # table of contents
    toc-location: left           # left | right | body
    number-sections: true        # section numbering
    number-depth: 4              # how deep to number (H4 and above)
    code-fold: true              # collapsible code chunks
    code-tools: true             # show/hide all code button
    self-contained: true         # single portable HTML file
    embed-resources: true        # embed fonts/images
    theme: flatly                # bootswatch theme name
    linkcolor: blue
    mainfont: Open Sans
    monofont: Fira Code
execute:
  echo: false      # hide code by default
  warning: false   # suppress warnings
  message: false   # suppress messages
  freeze: auto     # only re-render when source changes
```


# Useful Links

- [Quarto HTML format reference](https://quarto.org/docs/reference/formats/html.html)
- [Quarto callouts](https://quarto.org/docs/authoring/callouts.html)
- [Quarto figures](https://quarto.org/docs/authoring/figures.html)
- [Quarto code chunks](https://quarto.org/docs/reference/cells/cells-knitr.html)
- [Bootswatch themes](https://bootswatch.com/)
